Leadership across cultures and contexts: an exploratory analysis

### 2020-05-09 leadership_across_cultures_contexts.Rmd
Minor edits to variation section of discussion.

The distribution of text records by group context and subsistence strategy suggests that although leaders emerge within similar social contexts across populations, it occurs at variable frequencies. See Figure \@ref(fig:grouptype). Leadership in residential subgroups and kin groups was relatively more common in hunter-gatherer and pastoralist societies, populations more likely to experience high residential mobility [@wild_making_2019; @rigby_persistent_1985; @kelly_foraging_2013; @venkataraman_hunter-gatherer_2017]. Supracommunity political leadership was more common in horticultural and agricultural societies, which tend to rely heavily on cultivated land and maintain fixed, defensible territories. These patterns support perspectives suggesting residential mobility and increased costs of dispersal, subsistence intensification, and demographic transitions are associated with increasing complexity of sociopolitical organization among non-WEIRD populations [@johnson_evolution_1987; @steward_basin-plateau_1938; @powers_evolutionary_2014; @kantner_identifying_2010]. The only example of supracommunity leadership among hunter-gatherers was from the Andamanese hunter-gatherers by [@az02-001, p. 45], who discussed the roles of a "head chief" from a permanent residential location and their marginal influence over other, local residential leaders (see quote in Methods section). 
<!-- Thus, these results provide a fine-grained empirical view of putative cultural transitions in leadership systems. -->

@@

One possible interpretation of the negative relationships between political leadership at the community leadership or above and leadership in kin groups or residential groups is that as political structures become more institutionalized and economic systems transcend the household, leadership within households and residential groups diminishes. It is interesting, however, that even among nonindustrial agricultural populations which are characterized by leadership systems beyond the community level -- more similar to leadership systems of WEIRD populations -- the importance of leadership within residential and family groups persists. For example, among rural Korean populations a chairwomen of local organizations maintains important, possibly informal, leadership positions within the group and represents the group to governmental bodies. @aa01-062 [p. 173] describes that in these local 'clubs',

@@

Several leadership dimensions that are theoretically important were relatively rare in the ethnographic record. The leadership literature among nonhuman animals, for example, commonly focuses on group movement [@couzin_effective_2005], and "movement" was one of four leadership dimensions that @smith_leadership_2016 identified as comparable among small-scale human societies and nonhuman animal societies, yet movement or migration was identified as a leader function in only `r movement_pct`% of cultures, of which the vast majority (71%) were hunter-gatherers. Among the Blackfoot in the North American plains for example, @nf06-061 [p. 411] describes that, "if there was a dispute as to whether the camp should move north or south, the chief might present his arguments, gain support, and then say that he was going south. He did not order the others to follow, but he knew that they probably would go with him."

Several theories of leadership also emphasize the moral authority of leaders [e.g., @henrich_big_2015; @henrich_evolution_2001; @richerson_tribal_2009; @tietjen_moral_1985; @boggild_evolved_2016], yet leaders were seen as sources of moral authority or "fair" in only a minority of cultures (`r moral_pct`% and `r fairness_pct`% of cultures, respectively). The *Dominance* theory of leadership [e.g., @chapais_competence_2015; @cheng_dominance_2019; @cheng_pride_2010; @henrich_evolution_2001; @tiger_imperial_1971] emphasizes leaders' ability to instill fear, but in only `r feared_pct`% of cultures were leaders characterized as feared. Finally, a large literature on leadership in industrialized societies highlights the physical attractiveness of leaders [e.g., @sinclair_sexuality_1995; @altemeyer_sexual_1974; @march_leadership_2009], but leaders were characterized as attractive in only `r attractive_pct`% of the nonindustrial cultures sampled here. Clearly there is substantial diversity in many dimensions of leadership among non-WEIRD populations.

>>>>

The distribution of text records by group context and subsistence strategy suggests that although leaders emerge within similar social contexts across populations, it occurs at variable frequencies. See Figure \@ref(fig:grouptype). Leadership in residential subgroups and kin groups was relatively more common in hunter-gatherer and pastoralist societies, populations more likely to experience high residential mobility [@wild_making_2019; @rigby_persistent_1985; @kelly_foraging_2013; @venkataraman_hunter-gatherer_2017]. Supracommunity political leadership was more common in horticultural and agricultural societies, which tend to rely heavily on cultivated land and maintain fixed, defensible territories. These patterns support perspectives suggesting residential mobility and increased costs of dispersal, subsistence intensification, and demographic transitions are associated with increasing complexity of sociopolitical organization among non-WEIRD populations [@johnson_evolution_1987; @steward_basin-plateau_1938; @powers_evolutionary_2014; @kantner_identifying_2010]. The only example of supracommunity leadership among hunter-gatherers was from the Andamanese by @az02-001[p. 45], who discussed the roles of a "head chief" from a permanent residential location and their marginal influence over other, local residential leaders (see quote in Methods section). 
<!-- Thus, these results provide a fine-grained empirical view of putative cultural transitions in leadership systems. -->

@@

One possible interpretation of the negative relationships between political leadership at the community leadership or above and leadership in kin groups or residential groups is that as political structures become more institutionalized and economic systems transcend the household, leadership within households and residential groups diminishes. It is interesting, however, that even among nonindustrial agricultural populations which are characterized by political leadership systems beyond the community level -- more similar to political structures of WEIRD populations -- the importance of leadership within residential and family groups persists. For example, among rural Korean populations a chairwomen of local 'club' organizations maintains an important, although informal, leadership position within the group and represents the group to governmental bodies. @aa01-062 [p. 173] describes,

@@

Several theoretically important leadership dimensions were relatively rare in the ethnographic record. The nonhuman animal leadership literature, for example, commonly focuses on group movement [@couzin_effective_2005], and "movement" was one of four leadership dimensions that @smith_leadership_2016 identified as comparable among small-scale human societies and nonhuman animal societies, yet movement or migration was identified as a leader function in only `r movement_pct`% of cultures, of which the vast majority (71%) were hunter-gatherers. Among the Blackfoot in the North American plains for example, @nf06-061 [p. 411] describes that, "if there was a dispute as to whether the camp should move north or south, the chief might present his arguments, gain support, and then say that he was going south. He did not order the others to follow, but he knew that they probably would go with him."

Several theories of leadership also emphasize the moral authority of leaders [e.g., @henrich_big_2015; @henrich_evolution_2001; @richerson_tribal_2009; @tietjen_moral_1985; @boggild_evolved_2016], yet leaders were seen as sources of moral authority or "fair" in only a minority of cultures (`r moral_pct`% and `r fairness_pct`% of cultures, respectively). The dominance theory of leadership [e.g., @chapais_competence_2015; @cheng_dominance_2019; @cheng_pride_2010; @henrich_evolution_2001; @tiger_imperial_1971] emphasizes leaders' ability to instill fear, but in only `r feared_pct`% of cultures were leaders characterized as feared. Finally, a large literature on leadership in industrialized societies highlights the physical attractiveness of leaders [e.g., @sinclair_sexuality_1995; @altemeyer_sexual_1974; @march_leadership_2009], but leaders were characterized as attractive in only `r attractive_pct`% of the nonindustrial cultures sampled here. Clearly there is substantial diversity in many dimensions of leadership among non-WEIRD populations.
